Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-21 manuscript/ERL_global_C_review.Rmd
Update ERL_global_C_review.Rmd

*Review Results/ Synthesis.*  ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets. The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation. The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.

@@

(**check paragraph with latest data**) Average C cycles for forests <100 years old are presented in Figures 8-11.  Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs. 5-10, S1-S26; detailed below). Differences across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests (Figs. 6-7).

(**check paragraph with latest data**) ForC contained 14 flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**). Of these, 9 increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$). The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend. In terms of C stocks, 10 variables had sufficient data to test for age trends. Six of these—total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass—increased significantly with $log10[stand.age]$. The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.

@@

## Acknowledgements
All researchers whose data is included in ForC and this analysis.
Thanks to Norbert Kunert and [Helene's intern] for helpful input at an earlier phase.
A Smithsonian Scholarly Studies grant to KAT and HML. 
WLS grant to KAT.

>>>>

*Review Results/ Synthesis.*  ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets. The rate of C cycling generally increased from boreal to tropical regions in both mature and regrowth forests, whereas C stocks showed less directional variation. The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.

@@

Average C cycles for forests <100 years old are presented in Figures 8-11.  
Both C stocks and fluxes commonly displayed significant trends with stand age for within-biome analyses (Table 1, Figs. 6-11, S1-S26; detailed below).

ForC contained *14* flux variables with sufficient data for cross-biome analyses of age trends in regrowth forests (see Methods) (Fig. 6-7 and **S#- SI figures including plots for all variables**). 
Of these, *9* increased significantly with log10[stand.age]: $GPP$, $NPP$, $ANPP$, $ANPP_{foliage}$, $ANPP_{woody}$, $ANPP_{woody-stem}$, $BNPP$, $BNPP_{root-fine}$, $R_{eco}$, and net C sequestration ($NEP$). 
The remaining five--$ANPP_{woody-branch}$, $BNPP_{root-coarse}$, $R_{soil-het}$, and $R_{soil-het}$-–displayed no significant relationship to stand age, although all displayed a positive trend.

Differences in C fluxes across biomes typically paralleled those observed for mature forests, with C cycling generally most rapid in the tropics and slowest in boreal forests.  
The single exception was $ANPP_{stem}$, for which temperate broadleaf forests and temperate conifer forests of age >*~30* had slightly higher flux rates than tropical forests (*represented by a single chronosequence*).
Notably, the trend of tropical > temperate > boreal held for $NEP$ in regrowth forests, in contrast to the lack of biome differences in $NEP$ for mature forests (Fig. 6).

There were only *#* flux variables with sufficient data to test for biome x age interactions: $ANPP$, $ANPP_{woody}$, $ANPP_{stem}$, $ANPP_{litterfall}$, and $BNPP$. **(more could be added if age trends become significant after outliers are resolved)** 
For three of these ($ANPP$, $ANPP_{litterfall}$, $BNPP$), the increase in C flux with age was steepest increase in tropical forests, followed by temperate and then boreal forests (Figs S#). 
Similarly, $ANPP_{woody}$ displayed a steeper increase with age in temperate than boreal boreal forests (no tropical data for this variable).
In contrast, for $ANPP_{stem}$,  tropical and temperate broadleaf forests had the highest flux rates at young ages, but were surpassed by temperate conifer forests between ages 20 and 50 (Fig. S6).

(**this needs to be updated with latest data**) 
In terms of C stocks, *10* variables had sufficient data to test for age trends. 
*Six* of these—*total biomass, aboveground biomass, aboveground woody biomass, foliage biomass, root biomass, and coarse root biomass*— increased significantly with $log10[stand.age]$. 
*The remaining four displayed non-significant positive trends:  fine root biomass, total dead wood, standing dead wood, and organic layer.*
*(discuss rates of increase)*

@@

## Acknowledgements
All researchers whose data is included in ForC and this analysis.
Ian McGregor for help with the database.
Thanks to Norbert Kunert and [Helene's intern] for helpful input at an earlier phase.
A Smithsonian Scholarly Studies grant to KAT and HML. 
WLS grant to KAT.
